Target Speaker Extraction using Discrete Representations from Self-Supervised Models and Language Models \thanks{Identify applicable funding agency here. If none, delete this.}

### 2024-09-05 src/paper.tex
update

\author{\IEEEauthorblockN{1\textsuperscript{st} Beilong Tang}
\IEEEauthorblockA{\textit{Duke Kunshan University} \\
% \textit{name of organization (of Aff.)}\\
Kunshan, China \\
bt132@duke.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Zhan Jin}
\IEEEauthorblockA{\textit{Duke Kunshan University} \\
Kunshan, China \\
zhan.jin@whu.edu.cn}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Bang Zeng}
\IEEEauthorblockA{\textit{Duke Kunshan University} \\
% \textit{name of organization (of Aff.)}\\
Kunshan, China \\
bangzeng@whu.edu.cn}
\and
\IEEEauthorblockN{4\textsuperscript{th} Ming Li}
\IEEEauthorblockA{\textit{Duke Kunshan University} \\
% \textit{name of organization (of Aff.)}\\
Kunshan, China \\
ming.li369@dukekunshan.edu.cn}
}

@@

\subsubsection{Cross Attention}
One of the key factors in target speaker extraction is to 
inject the information of target speaker 
into the mixture. Spex+ \cite{spex_plus} jointly trains a speaker verification task module 
besides the separator, which might not be optimal for the separator task \cite{sef_net}. SEF-Net \cite{sef_net} has proposed an speaker embedding free way to inject the 
target speaker information into the mixture by using cross attention module, and it has 
achieved superior performance compared to other methods. Therefore, we apply 
the cross attention module to inject the reference embedding into the mixture. The details 
are in Fig.\ref{cross_attention}.
The cross attention module consists of a positional 
embedding layer, a cross-multiple head attention module, 
and a FiLM module. 
We use \(E_m\) as query and \(E_r\) as key and value for 
the attention module. The output of the cross-multiple head attention module \(E_{spk}\)
is passed together with \(E_m\) to the feature-wise linear modulation (FiLM) get the final output of the module. The output of 
FiLM is denoted as \(E_f = FiLM(E_m, E_{spk}) = \gamma E_{spk} \cdot E_m  + \beta E_{spk} \) where 
\(\gamma\) and \(\beta\) are learnable parameters denoting the scaling and shifting vectors 
respectively. The output is then fed to the language model for later modeling.  
\subsubsection{Language Modeling}
We use encoder-only language models containing multiple self-attention modules to model the injected 
embedding \(E_f\). Due to encoder-only style, the LM is able to learn from all the positions. 
Finally, 6 linear classifiers each with dimension \(K\) is used to produce the probability 
distribution of each layer respectively.
Cross-entropy loss is applied between 
the output tokens and the clean tokens through discretizing the 
ground truth clean audio .

@@

\subsubsection{Training sets}
We use the pretrained WavLM Large \cite{wavlm} and it is freezed during training. The Kmeans tokenizer is trained on the

>>>>

\author{\IEEEauthorblockN{1\textsuperscript{st} Beilong Tang}
\IEEEauthorblockA{\textit{Duke Kunshan University} \\
% \textit{name of organization (of Aff.)}\\
Kunshan, China \\
bt132@duke.edu}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Bang Zeng}
\IEEEauthorblockA{\textit{Duke Kunshan University} \\
% \textit{name of organization (of Aff.)}\\
Kunshan, China \\
bangzeng@whu.edu.cn}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Zhan Jin}
\IEEEauthorblockA{\textit{Duke Kunshan University} \\
Kunshan, China \\
zhan.jin@whu.edu.cn}
\and
\IEEEauthorblockN{4\textsuperscript{th} Ming Li}
\IEEEauthorblockA{\textit{Duke Kunshan University} \\
% \textit{name of organization (of Aff.)}\\
Kunshan, China \\
ming.li369@dukekunshan.edu.cn}
}

@@

\subsubsection{Cross Attention}
One of the key factors in target speaker extraction is to 
inject the information of target speaker 
into the mixture. Spex+ \cite{spex_plus} jointly trains a speaker verification task module 
besides the separator, which might not be optimal for the separator task \cite{sef_net}. SEF-Net \cite{sef_net} has proposed an speaker embedding free way to inject the 
target speaker information into the mixture by using cross attention module, and it has 
achieved superior performance compared to other methods. Therefore, we apply 
the cross attention module to inject the reference embedding into the mixture. The details 
are in Fig.\ref{cross_attention}.
The cross attention module consists of a positional 
embedding layer, a cross-multiple head attention module, 
and a FiLM module. 
We use \(E_m\) as query and \(E_r\) as key and value for 
the attention module. The output of the cross-multiple head attention module \(E_{spk}\)
is passed together with \(E_m\) to the feature-wise linear modulation (FiLM) get the final output of the module. The output of 
FiLM is denoted as \(E_f = FiLM(E_m, E_{spk}) = \gamma E_{spk} \cdot E_m  + \beta E_{spk} \) where 
\(\gamma\) and \(\beta\) are learnable parameters denoting the scaling and shifting vectors 
respectively. The output is then fed to the language model for later modeling.  
\subsubsection{Language Modeling}
We use encoder-only language models containing multiple self-attention modules to model the injected 
embedding \(E_f\). Due to encoder-only style, the LM is able to learn from all the positions. 
Finally, 6 linear classifiers each with dimension \(K\) is used to produce the probability 
distribution of each layer respectively.
Cross-entropy loss is applied between 
the output tokens and the clean tokens through discretizing the 
ground truth clean audio.

@@

\subsubsection{Training sets}
We use the publicly available Kmeans tokenizer and scalable 
HiFi-GAN decoder in \cite{speechbrain}. The Kmeans tokenizer
is trained on \texttt{train\_clean\_100}, \texttt{train\_clean\_360}, and \texttt{train\_other\_500} of
LibriSpeech \cite{librispeech}, and the scalable HiFi-GAN is trained on \texttt{train\_clean\_100} of 
LibriTTS \cite{libritts}. The modeling stage is trained on 
\texttt{train\_clean\_100} and \texttt{train\_clean\_360} of LibriSpeech. All training data are 
generated on the fly with relative SNR between 0 dB to 5 dB.
